Unifying Gompertzian growth with the communicable disease spreading paradigm

### 2022-04-27 main.tex
log-normal distribution implies CLT

Thus, a more parsimonious interpretation of the observations not reliant on a transmission network is that mortality was caused by a planetary perturbation, modeled as a random process, to which organisms gradually develop resistance at a geometric rate in the log-transformed domain \citep{boxenbaum2017hypotheses,neafsey1988gompertz}, which is the natural transformation for many processes in nature \citep{zhang1994log}. 
Under this model, the distribution of the abundance of $F(t)$ is log-normal, a result that can be obtained by directly solving the stochastic equation in \ref{eq:microscopicOriginal} \citep{skiadas2010exact,petroni2020gompertz}, or from thermodynamic principles \citep{sitaram1984statistical,gunasekaran1982lon}. 
Intuitively, this is seen by noting that the solution to the perturbation model in the log-domain (\ref{eq:microscopic}) is Gaussian in the variable $Z(t)$, thus suggesting a log-normal distribution of $F(t)$.

@@

In contrast, as shown in (\ref{eq:microscopic}), the Gompertz model is implied by a multiplicative stochastic perturbation and thus suggests correlated mortality growth, viz.
\begin{equation}
F(t) = \exp{\left[\frac{1}{N}\sum_i^N z_i(t)\right]} = \left[\prod_i^N f_i(t)\right]^{1/N},
\end{equation}
where the rightmost equality is due to the central limit theorem where $z_i$ is a random variable as in \ref{eq:microscopic} while $f_i$ is here deterministic. 
Under this model, correlation between entities is present at all orders in the original domain\footnote{It is illuminating to at this point compare with Gompertz' Law of Mortality, $\frac{d}{dt}\ln{(1 - F(t))} = -b$, for $t$ more than ~25 years, which yields a naturally uncorrelated macroscopic curve $1 - F(t) = \exp{(-bt)}$ \citep{shklovskii2005simple}. 
In our context, the uncorrelated feature emerges since the force of mortality is not a function of the growth itself, as we see in the text, but rather of time.}.

@@

\newpage
\section*{Supplement}

>>>>

Thus, a more parsimonious interpretation of the observations not reliant on a transmission network is that mortality was caused by a planetary perturbation, modeled as a random process, to which organisms gradually develop resistance at a geometric rate in the log-transformed domain \citep{boxenbaum2017hypotheses,neafsey1988gompertz}, which is the natural transformation for many processes in nature \citep{zhang1994log}. 
Under this model, the distribution of the abundance of $F(t)$ is log-normal, a result that can be obtained by directly solving the stochastic equation in \ref{eq:microscopicOriginal} \citep{skiadas2010exact,petroni2020gompertz}, or from thermodynamic principles \citep{sitaram1984statistical,gunasekaran1982lon}. 
Intuitively, this is seen by noting that the solution to the perturbation model in the log-domain (\ref{eq:microscopic}) is Gaussian in the variable $Z(t)$, thus suggesting a log-normal distribution of $F(t)$ and justifying a central limit theorem in this log-domain.

@@

In contrast, as shown in (\ref{eq:microscopic}), the Gompertz model is implied by a multiplicative stochastic perturbation and thus implies correlated mortality growth by application of the central limit theorem in the log-domain, viz.
\begin{equation}
F(t) = \exp{\left[\frac{1}{N}\sum_i^N z_i(t)\right]} = \left[\prod_i^N f_i(t)\right]^{1/N},
\end{equation}
where $z_i$ is a random variable as in \ref{eq:microscopic} while $f_i$ is here deterministic. 
Under this model, correlation between entities is present at all orders in the original domain\footnote{It is illuminating to at this point compare with Gompertz' Law of Mortality, $\frac{d}{dt}\ln{(1 - F(t))} = -b$, for $t$ more than ~25 years, which yields a naturally uncorrelated macroscopic curve $1 - F(t) = \exp{(-bt)}$ \citep{shklovskii2005simple}. 
In our context, the uncorrelated feature emerges since the force of mortality is not a function of the growth itself, as we see in the text, but rather of time.}.

@@

\section*{Code Availability}
The computer code required to produce the results are given in the following Github repository,
\begin{verbatim}
https://github.com/matzhaugen/GompertzManuscript
\end{verbatim}

\newpage

\section*{Supplement}
